---
title: "Assignment 6"
author: "Nick Andriese"
date: '2022-07-24'
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}

#install.packages("car","gvlma", "MASS", "leaps")
library(car)
library(gvlma)
library(MASS)
library(leaps)
```

We are going to look at the property prices for Orange County. From there we are going to try and predict the sales price by linear regression.

# Importing Data 

Bring in the data and make sure the data types are correct. If not, make the proper changes. The file is located within this project. *data/prop_prices_reduced.csv*

```{r}
library(readr)
prprice <- read_csv("prop_prices_reduced.csv")
View(prprice)
prprice$pool <- factor(prprice$pool)

```


# Plotting

Plot histograms for all variables. Additionally, add scatterplots for the relationships between all quantitative variables.

```{r}
library(car)
hist(prprice$sale_def)
hist(prprice$bed)
hist(prprice$bath)
hist(prprice$area_heated)
hist(prprice$area)
hist(prprice$dist_cbd)
hist(prprice$dist_lakes)
plot(prprice$pool)

scatterplotMatrix(prprice, spread = FALSE, lty.smooth = 2, main = "Scatterplot Matrix")


prprice1 <- prprice



#excess below

plot(prprice$bed, prprice$sale_def)
plot(prprice$bath, prprice$sale_def)
plot(prprice$area_heated, prprice$sale_def)
plot(prprice$area, prprice$sale_def)
plot(prprice$dist_cbd, prprice$sale_def)
plot(prprice$dist_lakes, prprice$sale_def)
plot(prprice$pool, prprice$sale_def)
plot(prprice$bed, prprice$bath)
plot(prprice$bed, prprice$area_heated)
plot(prprice$bed, prprice$area)
plot(prprice$bed, prprice$dist_cbd)
plot(prprice$bed, prprice$dist_lakes)
plot(prprice$bed, prprice$pool)
plot(prprice$bath, prprice$area_heated)
plot(prprice$bath, prprice$area)
plot(prprice$bath, prprice$dist_cbd)
plot(prprice$bath, prprice$dist_lakes)
plot(prprice$bath, prprice$pool)
plot(prprice$pool, prprice$area_heated)
plot(prprice$pool, prprice$area)
plot(prprice$pool, prprice$dist_cbd)
plot(prprice$pool, prprice$dist_lakes)
plot(prprice$area_heated, prprice$area)
plot(prprice$area_heated, prprice$dist_cbd)
plot(prprice$area_heated, prprice$dist_lakes)
plot(prprice$area, prprice$dist_cbd)
plot(prprice$area, prprice$dist_lakes)
plot(prprice$dist_cbd, prprice$dist_lakes)
```

# Summary Statistics

Provide basic summary statistics for univariate analysis. Also, provide the correlation between all the quantitative variables.

```{r}
#install.packages("psych")
library(psych)
summary(prprice)
describe(prprice)

cor(prprice[,unlist(lapply(prprice, is.numeric))])
```

# Regression Analysis

Run a regression with all the variables included. Print results of the regression.

```{r}
model1 <- lm(sale_def ~., data = prprice)

summary(model1)
```

Which of the variables tested significant at the 95% level? Looking at the results and answering outside of the chunk is sufficient.

***bed, bath, area_heated, area, dist_cbd all tested significant above the the 95% level

## Evaluating the model

As is, are any of the Gauss-Markov assumptions violated? If so, which ones? How can you fix the issues?

```{r}
#Regression diagnostics
opar <- par(no.readonly = TRUE)
par(mfrow =c(2, 2))
plot(model1)
###only seems like #37 is vast outlier, in trending downward on residuals

newmodel1 <- lm(sale_def ~., data = prprice[-c(37),])
plot(newmodel1)
### without 37 included

#Normality
qqPlot(model1, labels=FALSE, simulate = TRUE, main = "Q-Q Plot")
qqPlot(newmodel1, labels=FALSE, simulate = TRUE, main = "Q-Q Plot")
###appears to start curling around the tails, non-normal

#autocorrelated erros
durbinWatsonTest(model1)
durbinWatsonTest(newmodel1)
###not significant, no autocorrelation of errors

#what does this test???
crPlots(model1, one.page = TRUE, ask = FALSE)
crPlots(newmodel1, one.page = TRUE, ask = FALSE)

#Heteroskedasticity diagnostic:
ncvTest(model1)
ncvTest(newmodel1)
###both models indicate heteroskedasticity
###What should I do to correct for it? the power?
spreadLevelPlot(model1)
#heteroskedastic

#gvlma
library(gvlma)
gvmodel1 <- gvlma(model1)
summary(gvmodel1)

gbnewmodel1 <- gvlma(newmodel1)
summary(gbnewmodel1)

###all assumptions are violated: Global Stat, Skewness, Kurtosis, Link Function, Heteroskedasticity

#Outliers
outlierTest(model1)
outlierTest(newmodel1)

#Box Cox - Power Transformation
summary(powerTransform(prprice$sale_def/prprice$area))
#seems that both keeping it the same and taking the log of it will not be a good linear fit.

#BoxTidwell
boxTidwell(sale_def ~ bed +  area, data=prprice)
#tried this with all of the variables but kept throwing an error.

#Multicollinearity diagnostics
vif(model1)
##no concerns of multicollinearity

###model comparison
anova(model1, model2)
model2 <- lm(sale_def ~ bed + bath + area_heated + area + dist_cbd + pool, data=prprice)

newmodel2 <- lm(sale_def ~ bed + bath + area_heated + area + dist_cbd + pool, data=prprice[-c(37),])
summary(newmodel2)
anova(newmodel1,newmodel2)

summary(model1)
##
par(opar)
```

## New Model

Based off of your findings in the previous section, make changes to the variables, the functional form, etc.


###will try multiple attempts to fix assumptions

#normal model
model1 <- lm(sale_def ~., data = prprice)
summary(model1)
plot(model1)
summary(gvlma(model1))
### adjusted r^2 = .5438, no assumption satisfied on gvlma, dist_lake not significant, outliers of 7 and 37

#

#model without dist_lake
nolakemodel <- lm(sale_def ~ bed + bath + area + area_heated + dist_cbd + pool, data = prprice)
summary(nolakemodel)
plot(nolakemodel)
summary(gvlma(nolakemodel))
###adjusted r^2 = .5443, no assumptions satisfied

#

#model with dist_lake and outliers removed

nolakemodel1 <- lm(sale_def ~ bed + bath + area + area_heated + dist_cbd + pool, data = prprice[-c(7,37),])
summary(nolakemodel1)
plot(nolakemodel1)
summary(gvlma(nolakemodel1))
### adjusted r^2 = .6894, pool teetering on significance, only heteroskedasticity is acceptable on gvlma

#

#log of y
logmodel <- lm(log(sale_def) ~ bed + bath + area + area_heated + dist_cbd + pool, data = prprice)
summary(logmodel)
plot(logmodel)
summary(gvlma(logmodel))
###adjusted r^2 = .7393, all significant, only heteroskedasticity is acceptable on gvlma

#

#sqrt of y
sqrtmodel <- lm(sqrt(sale_def) ~ bed + bath + area + area_heated + dist_cbd + pool, data = prprice[-c(7),])
summary(sqrtmodel)
plot(sqrtmodel)
summary(gvlma(sqrtmodel))
###no assumptions satisfied but adjusted r^2 = .7895

#

#combine sales and area
SAmodel <- lm((sale_def/area) ~ bed + bath + area_heated + dist_cbd + dist_lakes + pool, data = prprice)
summary(SAmodel)
plot(SAmodel)
summary(gvlma(SAmodel))
###homoskedastic and link function satisfied, however adj r^2 = .2698, dist_cbd and pool not significant.

# log combined sales and area y
logSAmodel <- lm(log(sale_def/area) ~ bed + bath + area_heated + dist_cbd + (dist_lakes) + pool, data = prprice)
summary(logSAmodel)
plot(logSAmodel)
summary(gvlma(logSAmodel))
### adj r^2 = .2467, heteroskedasticity sat, dist_cbd and pool not significant


#log combined sales and area, log all IVs
logSAallmodel <- lm(log(sale_def/area) ~ log(bed) + log(bath) + log(area_heated) + log(dist_cbd) + log(dist_lakes) + (pool), data = prprice)
summary(logSAallmodel)
plot(logSAallmodel)
summary(gvlma(logSAallmodel))
### adj r^2 = .2704, pool not significant (not logged), heterosked and link function satisfied.

#sqrt combined sales and area
sqrtSAmodel <- lm(sqrt(sale_def/area) ~ bed + bath + area_heated + dist_cbd + (dist_lakes) + pool, data = prprice)
summary(sqrtSAmodel)
plot(sqrtSAmodel)
summary(gvlma(sqrtSAmodel))
### adj r^2 = .2697, dist_cbd and pool not significant, only heteroskedasticity satisfied

#boxcox-transformation
powermodel <- lm((sale_def^(-.33)) ~ bath + area_heated + dist_cbd + pool, data = prprice[-c(37,329),])
summary(powermodel)
plot(powermodel)
summary(gvlma(powermodel))

stepAIC(powermodel, direction = "backward")

hist(prprice$sale_def^-.33)
### highest I could get adj r^2 = .6678, no assumptions satisfied



############
### Researched this online to try and figure it out
boxCox(model1, family="yjPower", plotit = TRUE)
transale_def <- yjPower(prprice$sale_def, .22)



boxcoxmodel <- lm((sale_def^(.22)) ~ bed + bath + (area_heated) + (area) + (dist_cbd) + pool, data = prprice[-c(7),])
summary(boxcoxmodel)
plot(boxcoxmodel)
summary(gvlma(boxcoxmodel))
### adj r^2 of .7843, hetero and link function satisfied.

hist(prprice$sale_def^.22)

###


boxTidwell((sale_def) ~ bed + bath + area + area_heated + dist_cbd, other.x=~ pool, data = prprice[-c(7),])

BoxTidmodel <- lm(sale_def ~ I(bed^(-8.65)) + I(bath^(3.72)) + I(area^(10.09)) + I(area_heated^(20.75)) + dist_cbd, data = prprice[-c(7),])
summary(BoxTidmodel)
plot(BoxTidmodel)

### adj r^2 of .8866 but the plots are wonky as can be. Won't give the gvlma.

###############



#I can't get anything to budge with the full gvlma satisfied.


prprice1 <- data.frame(prprice)
prprice1$sareaheated <- (prprice$sale_def/prprice$area_heated)
prprice1$logsareaheated <- log(prprice$sale_def/prprice$area_heated)
prprice1$sqrtareaheated <- sqrt(prprice$sale_def/prprice$area_heated)
prprice1$salearea <- (prprice$sale_def/prprice$area)
hist(prprice1$salearea)
prprice1$logsalearea <- (log(prprice$sale_def/prprice$area))
hist(prprice1$logsalearea)
prprice1$sqrtsalearea <- (sqrt(prprice$sale_def/prprice$area))
hist(prprice1$sqrtsalearea)
prprice1$logsarea_heated <- (log(prprice$area_heated))
hist(prprice1$logsarea_heated)
prprice1$sqrtarea_heated <- (sqrt(prprice$area_heated))
hist(prprice1$sqrtarea_heated)
prprice1$logdist_lakes <- (log(prprice$dist_lakes))
hist(prprice1$logdist_lakes)
prprice1$sqrtdist_lakes<- (sqrt(prprice$dist_lakes))
hist(prprice1$sqrtdist_lakes)
prprice1$logbath <- (log(prprice$bath))
hist(prprice1$logbath)
prprice1$sqrtbath <- (sqrt(prprice$bath))
hist(prprice1$sqrtbath)
describe(prprice1)
describe(prprice)
#^^^^ this was for me to check kurtosis and skew, etc.
#################


```{r}


#######
####
##
#Conclusion, using boxcoxmodel
boxcoxmodel <- lm((sale_def^(.22)) ~ bed + bath + (area_heated) + (area) + (dist_cbd) + pool, data = prprice[-c(7),])
summary(boxcoxmodel)
plot(boxcoxmodel)
summary(gvlma(boxcoxmodel))
### adj r^2 of .7843, hetero and link function satisfied, normality not satisfied even after using the boxcox transformation. 
# After further study, violation of normality may not be significant if both linearity and homoskedasticity are present

# this was HELL

```



# Prediction

Based on the following inputs, predict the deflated sales price:

- 2 bed
- 2 bath
- area_heated = 1223
- area = 9750
- dist_cbd = 19368
- dist_lakes = 490
- no pool

```{r}
NewData <-data.frame("bed" = 2, "bath" =2, "area_heated" =1223, "area"=9750, "dist_cbd"=19368, "dist_lakes"=490, "pool"=factor(0))
x <- predict(boxcoxmodel, NewData)

PredictedCost <- (exp(1))^(log(x)/.22)
print(PredictedCost)

```